%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Medium Length Professional CV
% LaTeX Template
% Version 2.0 (8/5/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Rishi Shah 
%
% Important note:
% This template requires the resume.cls file to be in the same directory as the
% .tex file. The resume.cls file provides the resume style used for structuring the
% document.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[10pt]{resume} % Use the custom resume.cls style

\usepackage[left=0.5in,top=0.5in,right=0.5in,bottom=0.5in]{geometry} % Document margins
\newcommand{\tab}[1]{\hspace{.05\textwidth}\rlap{#1}}
\newcommand{\itab}[1]{\hspace{0em}\rlap{#1}}
\name{Thomas J. Brailey} % Your name
\address{8540 Costa Verde Blvd., CA 92122 \\ +1 (858) 610-2437} % Your address
%\address{123 Pleasant Lane \\ City, State 12345} % Your secondary addess (optional)
\address{thomasjbrailey@gmail.com \\ https://github.com/tjbrailey \\ www.linkedin.com/in/tjbrailey} % Your phone number and email

\begin{document}
	
	%----------------------------------------------------------------------------------------
	%	EDUCATION SECTION
	%----------------------------------------------------------------------------------------
	
	\begin{rSection}{Education}
		
		{\bf University of California, San Diego} \hfill {\em 2016 - 2020} 
		\\ Bachelor of Science, Data Analytics \hfill {\em summa cum laude}
		\\ Coursework: Spatial Data Analysis, Causal Inference, Applied Data Analysis, Game Theory, Research Design
		
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	%	WORK EXPERIENCE SECTION
	%----------------------------------------------------------------------------------------
	
	\begin{rSection}{Research Experience}
		
		\begin{rSubsection}{Jameel Poverty Action Lab (J-PAL)}{2020-Present}{Research Associate}{}
		\item Analyzing the effectiveness of the Integrated Child Development Service on key child outcomes.  
		\item Implementing rigorous econometric models and conduct exploratory data analysis.
		\item Editing manuscripts and preparing presentations for government leaders and other stakeholders.
		\end{rSubsection}
		
		\begin{rSubsection}{Center for the Study of African Political Economy}{2018-2020}{Research/Lab Assistant}{}
			\item Independently developed a large-N database of election statistics across all recognized states.
			\item Used R to combine and reshape multiple datasets, visualize the data, and perform statistical analysis.
			\item Manually inputted missing data from newspapers and reports to ensure completeness of data. 
		\end{rSubsection}
		
		\begin{rSubsection}{Center for Peace and Security Studies}{2018-2020}{Research Assistant}{}
			\item Used R to conduct quantitative analyses on large-N datasets.
			\item Generated graphs, tables, and figures in R and LaTeX for research papers and conference presentations.
			\item Wrote briefs on my data analysis and code and presented findings. 
			
		\end{rSubsection}
		
		%\begin{rSubsection}{Revelle College}{2018-2019}{Resident Assistant}{}
		%\item Fostered an inclusive community within Revelle through event programming and outreach.
		%\item Acted as a dependable resource for residents regarding academics, emotional support, and social justice. 
		%\item Received regular training on issues including autism awareness, sexual assault, and mental health issues. 
		%\end{rSubsection}
		
		\begin{rSubsection}{Scripps Institution of Oceanography}{2017-2018}{Government Relations Intern}{}
			\item Reviewed and researched congressional testimonies as they related to SIO projects.
			\item Compiled summaries and created databases of government policy and legislation.
			\item Offered and accepted paid position at end of internship.
		\end{rSubsection}
		
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	%	Publications
	%----------------------------------------------------------------------------------------
	
	\begin{rSection}{Publications}
		\itemsep -5pt
		\begin{rSubsection}{}{}{As Co-Author}{}
			Inbok Rhee and Thomas Brailey. Targeting the core or swing?: Campaign Strategy during the 2017 Kenyan Presidential Election. Work in Progress. March 2020.
		\end{rSubsection}
		\begin{rSubsection}{}{}{As Research Assistant}{}
			Erik Gartzke, Jon R. Lindsay, and J. Andres Gannon. After Deterrence: Explaining Conflict Short of War. Under Review. March 2020. \\
			J. Andres Gannon and Kerry Chavez. Cruisin’ for a Bruisin’: The Effects of US Cruise Missile Strikes Since the Gulf War. Work in Progress. March 2020.
		\end{rSubsection}
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	%	Conferences
	%----------------------------------------------------------------------------------------
	\bigskip
	\begin{rSection}{Conference Invites}
		\itemsep -5pt
		\item Midwestern Political Science Association (Chicago, 2020)
		\item International Studies Association (Honolulu, 2020)
		\item Southern California Research Conference (San Diego, 2019)
		\item Western Political Science Association (San Diego, 2019)
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	%	AWARDS
	%----------------------------------------------------------------------------------------
	
	\begin{rSection}{Awards}
		\itemsep -5pt
		\item Sanford Lakoff Award for the Most Outstanding Senior Thesis 2020.
		\item Pi Sigma Alpha 2017-2020. 
		\item Provost Honors 2016-2020. 
		\item Ernest C. Mort Leadership Excellence Award 2018. 
		\item Outstanding Resident 2017.
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	% Volunteer Experience
	%----------------------------------------------------------------------------------------
	\begin{rSection}{Volunteer Experience} \itemsep -5pt
		\item Co-founder of the Political Science Student Association at UCSD.
		\item Employment Assistant at the International Rescue Committee.
		\item Tutor for San Diego's Refugee Tutoring Program.
		%\item Campaign Intern and Data Analyst for Mike Levin’s congressional campaign.
		%\item Data assistant to the national coordinator of the Citizens Climate Lobby, UK. 
		%\item Member of the campus-wide Community Standards Board.
	\end{rSection}
	
	%----------------------------------------------------------------------------------------
	% Volunteer Experience
	%----------------------------------------------------------------------------------------
	\begin{rSection}{Skills} \itemsep -5pt
		\begin{rSubsection}{Spoken Languages}{}{}
			\item English, French.
		\end{rSubsection}
		\begin{rSubsection}{Technical Languages}{}{}
			\item Proficient: R, \LaTeX, STATA, ArcGIS, Git, Microsoft Office, SPSS.\\
			\item Beginner: Python, Tableau.
		\end{rSubsection}
	\end{rSection}
	
	\begin{rSection}{References}
		References available upon request. 
	\end{rSection}
	
\end{document}
