---
title: "Senior Thesis: Ontology"
author: "Tom Brailey"
date: "2020-01-05"
categories: ["Writing"]
tags: ["senior thesis", "segmental autonomy", "power sharing", "ontology"]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(magrittr)
library(ggplot2)
# Read in excel file 
files <- list.files("C:/Users/tbrai/Dropbox/github_private/SeniorThesis/data/",
                    "tjbrailey_psp_ontology.xlsx", 
                    full.names = TRUE
                    )
files <- files[2]

read_excel_allsheets <- function(filename) { 
  sheets <- readxl::excel_sheets(filename) 
  x <- lapply(sheets, function(X) readxl::read_excel(filename, sheet = X)) 
  names(x) <- sheets 
  x 
} 

out <- lapply(files, read_excel_allsheets)
basename(files)

psp_ont <- out[[1]]$Sheet1
psp_ont <- dplyr::as_data_frame(psp_ont)

# Select relevant variables
psp_ont_prep <- psp_ont %>%
  dplyr::select(child_0,
                child_1,
                provisions)
```

Thesis progress is slow but rewarding. I have a skeleton of my thesis and am working dilligently to fill out the rest of it. One thing that I have to show is my power-sharing ontology. It's a jargon-y word, I know, but it's a fun concept to implement in R.

# Ontology
Ontology is just a fancy word for family-tree. An ontology helps to visualize the relationship between upper-nodes and lower-nodes; parents and children. Imagine that I am node-0, my parents are node-1, and their parents (my grandparents) are the second node---its just a way of mapping hierarchical relationships. I use this idea to map out existing conceptions of power-sharing, where "concepts of power-sharing" is node-0, the broad form (i.e. consociational democracy, inclusive power-sharing, partition, power-dividng etc.) is the first node, and the second node denotes the components of power-sharing. That is, the mechanisms that make up the broad forms of power-sharing: segmental autonomy, mutual veto, coalition cabinets, and so on. During my literature review portion of the thesis, I made an Excel sheet of all the different conceptions of power-sharing. The code below is a nice example of how to visualize ontologies using this data. 

Let's look at the data:
```{r}
head(psp_ont_prep, 20)
```

Now we have to specify which column belongs where in the family-tree:  
```{r}
# Generate pathString as new column
psp_ont_prep$pathString <- paste("Power Sharing Provision",
                             psp_ont_prep$child_0,
                             psp_ont_prep$child_1,
                             psp_ont_prep$provisions,
                             sep = "|")
```

Now, using the awesome data.tree package, we format the data to work with equally awesome networkD3 package: 
```{r}
# Create list 
psp_tree <- data.tree::as.Node(psp_ont_prep, pathDelimiter = "|")

print(psp_tree, limit = 15)

psp_list <- data.tree::ToListExplicit(psp_tree, unname = TRUE)
```

Now we use networkD3 to create the visualization. There is a lot of flexibility in terms of formatting your plots, and you can mix networkD3 with htmlwidgets to add titles and captions to your plots.
```{r}
# Create visualization
psp_ontology_vis <- networkD3::diagonalNetwork(psp_list, 
                         fontSize = 14,
                         textColour = "black",
                         opacity = 100,
                         width = 1900,
                         height = 1400
                         )

psp_ontology_vis <- htmlwidgets::prependContent(psp_ontology_vis, htmltools::tags$h1("Full Mapping of Power-Sharing Conceptions"))

# Visualize
psp_ontology_vis
```

So here is the visualization of my literature review efforts for the senior thesis. As you can see, there are a lot of competing, and often overlapping, conceptions of power-sharing. I believe this visualization provides ample evidence of the need to disaggregate power-sharing, and study the "lower nodes" in order to establish its effectiveness.

Anyway, back to the code. Saving networkD3 plots is somewhat odd, so if anyone has any suggestions on alternatives, do let me know. Essentially, you save the plot as an .html file and then take a snapshot of the .html file using the webshot package. See below: 
```{r}
# Save visualizations
### networkD3::saveNetwork(psp_ontology_vis, file = paste0(here::here(), '/vis/psp_ontology_vis.html'))

### webshot::webshot("file:///C:/Users/yourusername/folder/folder/folder/vis/psp_ontology_vis.html", paste0(here::here(), "/vis/psp_ontology_vis.png"))
```

More to follow!